---
title: "data manipulation"
output: html_document
---

```{r}
library(tidyverse)
```

## load in the FAS Litters data 

```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor :: clean_names(litters_df)
```

## look at `select` function

choose some columns and not others 

```{r}
#only keep those two columns
select(litters_df, group, gd0_weight)
```
```{r}
#can also look at gd0_weight
select(litters_df, group, gd0_weight:gd_of_birth)
```
```{r}
#can specify what we want to get rid of 
select(litters_df, -litter_number)
```

renaming columns
```{r}
select(litters_df, GROUP = group, LITTer_NumBer = litter_number)
```
```{r}
#dedicating function in dplyr
rename(litters_df, GROUp = group, LITTer_NumBer = litter_number)
```

select helpers -- want to use functions that help me select columns that i know i want to keep based on some criteria 
```{r}
select(litters_df, starts_with("gd"))
```

```{r}
#moves litter_number to front and keep everything else 
select(litters_df, litter_number, everything())
```

```{r}
#does the same thing -- select is the key function but there could be other functions that do the same thing
relocate(litters_df, litter_number)
```

# looking at `filter` function 

intended to remove or keep rows that we are interested in retaining for our analyses 
  - needs to be given a condition 

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
#can filter in the opposite direction
filter(litters_df, gd0_weight >= 22)
```
logical operators on P1805 website 
```{r}
## two == is a test for equality
filter(litters_df, gd_of_birth == 20)
```

```{r}
# ! gives us all the  variables when gd_of_weight is not equal to 20
  # this way is preferred when you have more complicated things 
## can also use != to mean not equal to
filter(litters_df, !(gd_of_birth == 20))
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20) 
```
```{r}
## when you want non-numeric numbers 
filter(litters_df, group == "Mod8")
```

```{r}
## want two variables -- where group is either of con7 or mod8 option 
filter(litters_df, group %in% c("Con7","Mod8"))
```



## look at `mutate` function
    - able to create some additional variable or modify an existing variable 
  
```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```
```{r}
## can also mutate by adding multiple variables at the same time 
## create new variable (first code), modify an existing variable (second code)
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```

## looking at `arrange` function 
goal is to put things in order -- don't use that often

```{r}
arrange(litters_df, pups_born_alive)
```

## pipe operation `%>%` 

DO NOT DO THIS (NESTING)
```{r}
litters_df_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_df_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

instead, USE PIPE OPERATOR INSTEAD (shift, command, m) 
  - easier to read and take a look at 
```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```

